---
title: "Survival Rate - Multi parameter analysis"
format: html
editor: visual
---

## Loading Libraries

```{r}
library('tidyverse')
library('survival')
library('here')
library("viridis")
```

## Loading Data

```{r}
patients_sample_gene_panel <- read_delim(here("data", "patient_sample_gene_panel"), delim = "\t", col_names = TRUE)

```

## Data Wrangling

Selecting our columns of interest (originally form clinical patient and clinical sample files).

```{r}
patients_sample <- patients_sample_gene_panel |>  
  select(c(SEX, OS_MONTHS, OS_STATUS, AGE_GROUP, DRUG_TYPE, CANCER_TYPE, TMB_NONSYNONYMOUS, AGE_AT_SEQ_REPORT))


```

```{r}
#Comparing the number of mutation, and the drug given, in the case of Glioma cancer
patients_sample <- patients_sample |> 
  mutate(TMB_group = case_when(TMB_NONSYNONYMOUS<5 ~'Low TMB',
                               5<=TMB_NONSYNONYMOUS & TMB_NONSYNONYMOUS<20 ~'Intermediate TMB',
                               TMB_NONSYNONYMOUS>=20 ~'High TMB'
                               ) |> 
           factor(levels = c('Low TMB', 'Intermediate TMB', 'High TMB'),
                  ordered = TRUE
                  ),
         AGE_GROUP = factor(AGE_GROUP,
                            levels = c("<30", "31-50", "50-60", "61-70", ">71"),
                            ordered = TRUE
                            )
         )

patients_sample |> 
  ggplot(mapping = aes(x = TMB_group,
                       fill = DRUG_TYPE)
         )+
  geom_bar()+
  labs(tile = 'Bar plot of Drug given againts different severity of TMB (Tumor Mutational Burden', x = 'TMB Severity')
```

## Analysis by Cancer Type

### Survival rate analysis

Does the cancer type affects the survival of patients?

```{r}

KM_f <- survfit(Surv(OS_MONTHS, OS_STATUS) ~CANCER_TYPE, data = patients_sample)

#convert to a tibble

print(KM_f$strata)
tibble_KM_f <- tibble(
  time = KM_f$time,
  survival = KM_f$surv,
  strata = rep(names(KM_f$strata), KM_f$strata))

tibble_KM_f <- tibble_KM_f %>%
  mutate(strata = str_remove(strata, "CANCER_TYPE="))

ggplot(tibble_KM_f,
       mapping = aes(x = time, 
                     y = survival, 
                     color = strata))+
  geom_step()+
  labs(title = 'Survival Curve of the patients depending on their Cancer type', x = 'Time (in months)', color = 'Cancer Type')+
  scale_fill_viridis(option = "viridis",
                     discrete = TRUE) 

  
```

#### Further analysis of patients based on their cancer type

```{r}
ggplot(patients_sample,
       mapping = aes(x = AGE_AT_SEQ_REPORT, 
                     fill = CANCER_TYPE))+
  geom_bar()+
  labs(title = 'Reaprtition off cancers at the different diagnosis ages', x ='Age at Diagnosis')+
  scale_fill_viridis(option = "viridis",
                     discrete = TRUE) 
  
```

```{r}
ggplot(patients_sample,
       mapping = aes(x = factor(TMB_group, order = TRUE), 
                     fill = CANCER_TYPE))+
  geom_bar()+
  labs(title = 'Proportion of cancer types depending on their TMB severity', x = 'TMB severity')+
  scale_fill_viridis(option = "viridis",
                     discrete = TRUE) 
```

```{r}
ggplot(patients_sample,
       mapping = aes(x = CANCER_TYPE, 
                     fill = SEX))+
  geom_bar()+
  labs(title = 'Proportion of males and females for each cancer type', x ='ancer Type')+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_viridis(option = "viridis",
                     discrete = TRUE, 
                     alpha =0.7) 
```

```{r}
ggplot(patients_sample,
       mapping = aes(x = CANCER_TYPE, 
                     fill = DRUG_TYPE))+
  geom_bar()+
  labs(title = 'Proportion of the type of drug given for each Cancer Type', x = 'cancer Type')+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_viridis(option = "viridis",
                     discrete = TRUE, 
                     alpha = 0.7) 
```

the impact of TMB on all cancer survival rate:

Does the tumor mutational burden (liked to the number of mutation) affect the survival and thus is an indicator of the severity of the cancer?

```{r}
KM_f <- survfit(Surv(OS_MONTHS, OS_STATUS) ~TMB_group, data = patients_sample)

#convert to a tibble

tibble_KM_f <- tibble(
  time = KM_f$time,
  survival = KM_f$surv,
  strata = rep(names(KM_f$strata),KM_f$strata))

tibble_KM_f <- tibble_KM_f %>%
  mutate(strata = str_remove(strata, "TMB_group="))

ggplot(tibble_KM_f,
       mapping = aes(x = time, 
                     y = survival, 
                     color = strata))+
  geom_step()+
  labs(title = 'Survival Curve of the patients depending on their TMB', x = 'Time (in months)', color = 'TMB severity', )+
  scale_fill_viridis(option = "viridis",
                     discrete = TRUE)
```

Higher TMB leads to higher survival.

What about in Glioma?

## Diving into Glioma

```{r}
patients_sample_glioma <- patients_sample |> 
  filter(CANCER_TYPE == "Glioma")
```

### Survival Analysis stratified by Tumor Mutational Burden

```{r}

KM_f <- survfit(Surv(OS_MONTHS, OS_STATUS) ~TMB_group, data = patients_sample_glioma)

#convert to a tibble

tibble_KM_f <- tibble(
  time = KM_f$time,
  survival = KM_f$surv,
  strata = rep(names(KM_f$strata),KM_f$strata))

tibble_KM_f <- tibble_KM_f %>%
  mutate(strata = str_remove(strata, "TMB_group="))

ggplot(tibble_KM_f,
       mapping = aes(x = time, 
                     y = survival, 
                     color = strata))+
  geom_step()+
  labs(title = 'Survival Curve of the patients depending on their TMB', x = 'Time (in months)',color = 'TMB severity', )+
  scale_fill_viridis(option = "viridis",
                     discrete = TRUE) 
```

There is maybe not enough High TMB to make a concrete conclusion here, the survival curve of high TMB could be biased.

Lets look more into it, what is the distribution of TMB in Glioma?

#### Further analysis of Glioma cancer

```{r}
#looking at the repartition of TMBs in the different sections
ggplot(patients_sample_glioma, 
       mapping = aes(x = TMB_group))+
  geom_bar()+
  labs(title = 'Repartition of TMB severity among patients with Glioma', x ='TMB severity')
```

```{r}
ggplot(patients_sample_glioma, 
       mapping = aes(x = (TMB_group), 
                     fill = DRUG_TYPE))+
  geom_bar()+
  labs(title = 'Drug given to Glioma patients depending on their TMB severity', x = 'TMB severity')+
  scale_fill_viridis(option = "viridis",
                     discrete = TRUE, 
                     alpha = 0.7) 
```

```{r}
KM_f <- survfit(Surv(OS_MONTHS, OS_STATUS) ~Sex, data = patients_sample_glioma)

#convert to a tibble

tibble_KM_f <- tibble(
  time = KM_f$time,
  survival = KM_f$surv,
  strata = rep(names(KM_f$strata),KM_f$strata))

tibble_KM_f <- tibble_KM_f %>%
  mutate(strata = str_remove(strata, "TMB_group="))

ggplot(tibble_KM_f,
       mapping = aes(x = time, 
                     y = survival, 
                     color = strata))+
  geom_step()+
  labs(title = 'Survival Curve of the patients depending on their TMB', x = 'Time (in months)', color = 'Gender')+
  scale_fill_viridis(option = "viridis",
                     discrete = TRUE) 
```
